---
title: "Developing Metrics for Location Scheduling"
author: "Amber Medina, FLIC Votes, Alachua County"
output:
  html_notebook: default
  pdf_document: default
---
``` {r echo = FALSE}
library(tidyverse)
library(googlesheets)
library(leaflet)
library(sqldf)
```
#{.tabset}
######Data does not reflect actual canvassing shifts. Names, collection numbers, and locations have been altered to protect privacy.

##Background
Which locations, independent of canvasser, are truly hot spots? 
 	
To answer these questions, we need every bit of information about **where** canvassing takes place, **how long** we canvassed, **how many forms** were collected, if those forms were **complete** or **incomplete**, and what an **average form collection** rate might be for that location. 

Starting with the location info, we need to ask Blocks for all of the **locations, addresses, and territories** and *all* shift information tied to locations. This information is not simultaneously available, but can be exported for you from your Shifts report and Locations table.

##Data Dictionary
The `locations.csv` will contain:  
  
``$name`` -Staging site name as submitted by QC  
``$street_address`` - QC Staff should have entered this as ## STREET 1 but watch for incomplete street names. This will affect geocoding!  
``$city`` -Same as above. Check spelling!  
``$state`` - All locations will be in FLORIDA.  
``$zipcode`` -Check that zip is 5 digits beginning with '3'  
``$county`` -Check spelling and hyphenation for 'Miami-Dade'  
``$latitude``  
``$longitude``  
  
  
The `shifts.csv` will contain:  
  
``$shift_id`` -Unique Identifier from Blocks  
``$packet_name`` - File submitted by QC corresponding to packet  
``$canvasser_first`` - Canvasser staged at that location  
``$location_name`` -Name of location as submitted by QC  
``$shift_date`` -Date of clock-in  
``$field_start`` -Field start time  
``$field_end`` -Field end time  
``$forms_collected`` -Number of complete forms in packet  
``$turf_name`` - Office name  


##1.Tidying Up
```{r echo = TRUE, warning = FALSE}
library(googlesheets)
#Import shift information
shifts <- gs_title("Shifts_TotalForms_20180621.csv")

shifts <- gs_read(shifts)

#Import locations table
locations <- gs_title("FLIC_Locations_06222018.csv")

locations <- gs_read(locations)
```

```{r echo=FALSE}
#Preview
head(shifts)
head(locations)
```
##2. Shifts
```{r echo = TRUE}
#Preview columns
colnames(shifts)

#Remove Columns we don't need
shifts <- select(shifts, -shift_id, -field_start_time, -field_date, -packet_name, - shift_start, -shift_end)
	
#New
head(shifts)
```

###Calculate Forms Per Hour
$shift\;length = (start - end)\;\div\;3600s$
```{r echo = TRUE}
#Create a new column populated with the time difference between start and  end. The result is in seconds, so we divide by 3600 to convert to hours.    
shifts <- shifts %>% mutate(shift_length = field_end - field_start) %>%
	mutate(shift_length = shift_length/3600)  
    
#Coercing shift_length from difftime to numeric.  
shifts$shift_length <- as.numeric(shifts$shift_length)  
    
#Look at shift_length! There should be no negative values! These were introduced when a QC person put in the wrong shift start or end time. We will have to take out all of these erroneous entries.
negatives <- which(shifts$shift_length <= 0)  
shifts <- shifts[-negatives, ]  
    
#Shifts now contains a column of numeric duration in hours, and we're ready to look at average form collection.  
head(shifts$shift_length)  
```

$forms\;per\;hour= \;forms\;collected\;\div\;shift\;length$
```{r}
# Create a new column populated with the forms collected per hour.
shifts <- shifts %>% mutate(formsperhour = forms_collected/shift_length)

#New column containing the ratio of forms per hour
head(shifts$formsperhour)

```


##4. Locations
  `locations.csv` contains all of the public canvassing locations and their addresses. You can geocode this using this [tutorial](https://andrewbtran.github.io/NICAR/2017/maps/mapping-census-data.html#geolocating_addresses_in_r)
  
```{r echo = TRUE}
#Taking out all of the empty columns.
locations <-select(locations, -9, -10, -11, -12, -13, -14, -15, -16, -17, -18, -19, -20, -21, -22)

head(locations)
```
  For tables with `latitude` and `longitude` we need to append those values to the shifts data frame matching on the location name. I don't want to mess with the `sqldf` package, so I'm going to use the handy `left_join()` in `dplyr`. 
  
```{sql echo = TRUE, eval = FALSE}
--What we would have had in SQL syntax:
SELECT * FROM shifts AS s
JOIN locations AS l
ON s.location_name = l.location_name
```

```{r echo = TRUE}
#What is wonderful about R:
FLIC_FPH <- left_join( x = shifts, y = locations, by = "location_name") 
```
##5.First map
```{r eval = TRUE}
#package
library(leaflet)
library(colorspace)

take <- which(is.na(FLIC_FPH$latitude))
FLIC_FPH <- FLIC_FPH[-take, ]

#summary
summary(FLIC_FPH$formsperhour)

#create bins
FLIC_FPH$fph_range <- cut(FLIC_FPH$formsperhour, 
										 breaks = c(0.8, 1, 2, 38),
										 right = TRUE,
										 labels = c("Bad [0.8-1)", "Good [1-2)", "Best [2-38)"))

#define colors
pal <- colorFactor( palette = c("#A01606", "#FFC300", "#0D982F"), domain = FLIC_FPH$fph_range)


m <- leaflet(FLIC_FPH) %>% 
	addTiles() %>% 
	setView(-82.36376, 29.61666, zoom = 12) %>% 
	addCircleMarkers(
		lat = ~latitude,
		lng = ~longitude,
		color = ~pal(fph_range),
		radius = 2.5,
		opacity = 0.75,
		label = paste("Site:", FLIC_FPH$location_name)) %>%
	addLegend( position = "topleft", pal = pal, values = FLIC_FPH$fph_range, title = "Form Collection Rate")

m
```
##6. Grouped by Location

